Psychophysics of musical elements in the discrete-time representation of sound

### 2013-10-07 article/article.tex
feito ateh repeticao e variacao

Strong units (accents) fall in the `head' of bars, the downbeats, of which the first in
every bar is the greatest. The
head of a unity is the first part of the subdivision. In binary divisions (2, 4
and 8, for example), the unities originally strong turn into weak (não sei se isso ficou claro)
(e.g.\ division in 4 is: strong, weak, strong medium, weak). In ternary divisions
(3, 6 and 9) two weak unities succeeds the downbeat (e.g.\ division in 3 is:
strong, weak, weak)\footnote{Division in 6 is considered compound but can also
occur as a binary division. A binary division which suffers a ternary division
results in two units divided in tree units each: strong (subdivided in strong,
weak, weak) and weak (subdivided in strong, weak, weak). Another way to perform
the division in 6 is applying ternary division that suffers a binary division,
resulting in: a strong unit (subdivided in strong and weak) and two weak units
(subdivided in strong and weak for each).}

The accent in the weak beat is known as the backbeat, whereas a note starting in a weak
beat which duration coming across the strong beat is known as syncope.

Notes can occur inside and outside of these \emph{'music metric'} divisions. In
most behaved cases, notes occur exactly on these divisions, with broader
incidence on attacks of strong beats.
In extreme cases, the time metric cannot be perceived~\cite{Roederer}. 
Small grid variations help to build musical interpretation and differences
between styles~\cite{Cook}.

% ficou um tanto confuso:

Being pulse the grouping level $j=0$, level $j=-1$ the first pulse subdivision,
level $j=1$ the first pulse agglomeration and so on. In this way, $P_i^j$ is the
$i$-th unit of pulses at grouping level $j$: $P^0_{10}$ is the tenth pulse,
$P^{1}_3$ is the third pulse grouping unit (and, possibly, the third measure),
$P^{-1}_2$ is the second unit of pulse subdivision.

Special attention should be given to the limits of $j$: pulse divisions are
durations sensible as rhythm. Furthermore, the pulse joints sums, at its
maximum, a music or a cohesive set of musics. In other words, a duration given
by $P^{min(j)}_i$, $\forall \; i$, is greater than $50 ms$ and the durations
summed together $\sum_{\forall i}P^{\text{max}(j)}_i$ are less than a few
minutes or at most, a few hours.

Each level $j$ has some indexes $i$. A index $i$ always has three different
values (or multiple of three) that occur in a perfect relation. When $i$ has only
multiple of two, four or eight different values, than an imperfect relation occurs,
as shown in Figure~\ref{fig:pulsoSubAgl}.

Any unit (note) of a given musical sequence which has a time metric can be unequivocally
specified as:

@@

\noindent where $j_k$ is the grouping level and$i_k$ is the unit order itself.

As example, consider $P^{-1,0,1}_{3,2,2}$ as the third subdivision $P^{-1}_3$ of the
second pulse $P^0_2$ and of the second pulse group $P^1_2$. Each unit or unit set
$P_i^j$ can be associated with a sequence of temporal samples $T_i$ that forms a
music note.

The music piece \emph{Poli Hit Mia} explorers different metrics (avaleble in Appendix~\ref{ap:poli} and as part of \massa).

\subsection{Repetition and variation: motifs and large units}\label{subsec:motivos}

>>>>

Strong units (accents) fall in the `head' of bars, the downbeats, of which the first in
every bar is the greatest. The
head of a unit is the first part of the subdivision. In binary divisions (2, 4
and 8), strong units alternate with weak units
(e.g.\ division in 4 is: strong, weak, average strong, weak). In ternary divisions
(3, 6 and 9) two weak units succeeds the downbeat (e.g.\ division in 3 is:
strong, weak, weak). Division in 6 is considered compound but can also
occur as a binary division. Binary division units which suffers a ternary division
yields two units divided in tree units each: strong (subdivided in strong,
weak, weak) and weak (subdivided in strong, weak, weak). Another way to perform
the division in 6 is ternary division units that divides as binary,
resulting in: a strong unit (subdivided in strong and weak) and two weak units
(subdivided in strong and weak for each).

The accent in the weak beat is the backbeat, whereas a note starting in a weak
beat persisting across the strong beat is the syncope.

Notes can occur inside and outside of these \emph{'musical metric'} divisions. In
most well behaved cases, notes occur exactly on these divisions, with 
greater incidence strong beat attacks.
In extreme cases, time metric cannot be perceived~\cite{Roederer}. 
Small variations along the grid are crucial for musical interpretation and
different styles~\cite{Cook}.

Be the pulse the grouping level $j=0$, level $j=-1$ is the first pulse subdivision,
level $j=1$ is the first pulse agglomeration and so on. In this way, $P_i^j$ is the
$i$-th unit at grouping level $j$: $P^0_{10}$ is the tenth pulse,
$P^{1}_3$ is the third grouped unit (and, possibly, the third measure),
$P^{-1}_2$ is the second part of pulse subdivision.

The limits of $j$ are of special interest: pulse divisions are
durations perceivable as rhythm; furthermore, the pulses sum, at its
maximum, a music or a cohesive set of musics. In other words, a duration given
by $P^{min(j)}_i$, $\forall \; i$, is greater than $50 ms$ and the durations
summed together $\sum_{\forall i}P^{\text{max}(j)}_i$ are less than a few
minutes or, at most, a few hours.

Each level $j$ has some sections $i$. When $i$ has three different
values (or multiple of three) there is a perfect relation. When $i$ has only
two, four or eight different values, than there is an imperfect relation,
as shown in figure~\ref{fig:pulsoSubAgl}.

Any unit (note) of a given musical sequence with a time metric can be unequivocally
specified as:

@@

\noindent where $j_k$ is the grouping level and$i_k$ is the unit itself.

As an example, consider $P^{-1,0,1}_{3,2,2}$ as the third subdivision $P^{-1}_3$ of the
second pulse $P^0_2$ and of the second pulse group $P^1_2$ (possibly second measure).
Each unit $P_i^j$ can be associated with a sequence of temporal samples $T_i$ that comprehends a
note.

The music piece \emph{Poli Hit Mia} explorers different metrics (available as part of \massa\ toolbox).

\subsection{Repetition and variation: motifs and larger units}\label{subsec:motivos}
